---
title: "Final Project"
author: "Laia Esquerrà"
date: "June 30, 2017"
output: html_document
---

Correu: `bartekskorulski@gmail.com`

#Exploratory Analysis

```{r, include=FALSE, warning=FALSE}
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)
```

Given the data from *Instacart* and the previous analysis we made, we propose here some further approaches to it. 

First of all we'll analyse the pattern of the last minute purchases. Forehand, it would make sense that these follow a different pattern than general purchases, being this purchases made during the latter hours of the day directly related to some fast dinner arrangement or similar.

On the other hand, we'll analyse the last products added to a shopping cart. A first hypothesis in this case could be that those last items are usually not planned forehand, i.e., they have impulsively been added responding to last minute whims or as a result of the given recommendations from the app.

Finally, we'll see how regular consumers behave in relation to the rest.

##Last minute purchases

As we've previously said, we expect that this last minute purchases, done at late night hours might show a differential behaviour from the rest of the orders.

To analyse this, we'll select the most ordered products for puchases made after 8pm. 

```{r}
last_hour_txt <- "
SELECT ab.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM (SELECT order_id 
          FROM orders_tbl
          WHERE order_hour_of_day > 20) a
      LEFT JOIN order_products__prior_tbl b
      ON a.order_id = b.order_id
    WHERE product_id IS NOT NULL
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) ab
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) c
ON ab.product_id = c.product_id
"

(last_hour <-
  dbGetQuery(sc, last_hour_txt))
```

```{r}
last_hour %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='yellowgreen') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("Product") +
  ylab(paste("Popularity")) +
  theme_minimal()
```

On the contrary, these purchases same to follow the exact same pattern than the rest, being bananas clearly the most ordered product.

Note that this data comes from an app where customers select groceries online from various retailers and these are delivered by a personal shopper. Therefore, the fact that an order is taken in the evening shows no differential pattern, thus this purchases will probably be delivered the next day. It doesn't really correspond to a last minute purchase but to customers who make their orders at night when at home.

##Last added products

Which are the last added products in a basket? Do they follow a pattern or do they appear to be random and thus, probably unnecessary, resulting from a last minute impulse?

To analyse this we will select those products which, for each order of the first 50 orders, have been the latter one added.

```{r}
last_prods_txt <- "
SELECT ab.product_id, product_name, COUNT(1) as n
FROM
  (SELECT b.*
  FROM
    (SELECT *
    FROM (SELECT order_id
          ,   LAST_VALUE(add_to_cart_order)  OVER(partition by order_id ORDER BY order_id) AS nmax
          FROM order_products__prior_tbl
          WHERE order_id <= 50) a
    GROUP BY order_id, nmax
    ORDER BY order_id) aa
      LEFT JOIN order_products__prior_tbl AS b
      ON aa.nmax=b.add_to_cart_order
  WHERE aa.order_id=b.order_id
  ORDER BY order_id) ab LEFT JOIN products_tbl c
    ON ab.product_id = c.product_id
GROUP BY ab.product_id, product_name
ORDER BY n DESC
"

(last_prods <-
  dbGetQuery(sc, last_prods_txt))
```

```{r}
last_prods %>%
  ggplot(
    aes(reorder(product_id, n, function(x) x), 
        n)) +
  geom_bar(stat="identity", fill='pink') +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(label=scales::comma) +
  xlab("Products") +
  ylab(paste("Popularity")) +
  theme_minimal()
```

As we can see there is no particular product which tends to be added at the end, we can analyse though if they all come from some particular department.

**Departments**

Instead of counting which particular products have been selected, here we'll focus on the department they come from.

```{r}
last_prods_dpt_txt <- "
SELECT department_id, COUNT(1) as n
FROM
  (SELECT b.*
  FROM
    (SELECT *
    FROM (SELECT order_id
          ,   LAST_VALUE(add_to_cart_order)  OVER(partition by order_id ORDER BY order_id) AS nmax
          FROM order_products__prior_tbl
          WHERE order_id <= 50) a
    GROUP BY order_id, nmax
    ORDER BY order_id) aa
      LEFT JOIN order_products__prior_tbl AS b
      ON aa.nmax=b.add_to_cart_order
  WHERE aa.order_id=b.order_id
  ORDER BY order_id) ab LEFT JOIN products_tbl c
    ON ab.product_id = c.product_id
GROUP BY department_id
ORDER BY n DESC
"

(last_prods_dpt <-
  dbGetQuery(sc, last_prods_dpt_txt))
```

```{r}
last_prods_dpt %>%
  ggplot(aes(reorder(department_id, n, function(x) x), 
        n)) +
    geom_bar(stat="identity", fill='turquoise') +
    scale_y_continuous(label=scales::comma) +
    coord_flip() + 
    xlab("Order") +
    ylab("Number of products") +
    theme_minimal()
```

Now we can see that, the last added products in a cart come mainly from the 4th department, as well as many from the 19th. These correspondingly being *Produce (Fruits & Vegetables)* and *Snacks*. The first is by far the most common department, not in the amount of different products they have, but the one from which the most ordered products are from. Therefore, it is not uncommon for it to also be the most relevant here. Our interest falls on the latter though. As we suggested, many products which are last added to a cart are whims.

**Reordered**

Finally, we can analyse if this products are less reordered than the rest of products in the basket. Therefore, we'll calculate the ratio of reordered products which have been added last to an order and we'll compare it to the general ratio for these users.

```{r}
last_prods_reord_txt <- "
  SELECT reordered, COUNT(1) as n
  FROM
    (SELECT *
    FROM (SELECT order_id
          ,   LAST_VALUE(add_to_cart_order)  OVER(partition by order_id ORDER BY order_id) AS nmax
          FROM order_products__prior_tbl
          WHERE order_id <= 50) a
    GROUP BY order_id, nmax
    ORDER BY order_id) aa
      LEFT JOIN order_products__prior_tbl AS b
      ON aa.nmax=b.add_to_cart_order
  WHERE aa.order_id=b.order_id
  GROUP BY reordered
  ORDER BY n DESC
"

(last_prods_reord <-
  dbGetQuery(sc, last_prods_reord_txt))

#ratio
round(last_prods_reord[2,2]/sum(last_prods_reord[1:2,2]),3)
```

```{r, warning=FALSE}
order_products__prior %>%
  filter(order_id <= 50) %>%
  select(reordered) %>%
  group_by(reordered) %>%
  summarise(n =n())) %>%
  collect
```

```{r}
#ratio
228/(228+176)
```

As we can see, the latter added products are less frequently reordered products from previous purchases, which would be consistent with our first hypothesis that they are usually not planned forehand but instead, have been added more impulsively.

##Regular consumers

In this case we are interested in analysing if a person who buys everyday repeats products, and if their consecutive purchases are completely different or clearly structured.

Both cases would make sense, as a person who buys regularly, in the most extreme case, everyday for the correponding meals; will have very different orders from one day to another if he/she's keeping a balanced diet. At the same time, his/her orders will become more systematic and will present a clear structure, which will get plotted on the order reordered products are ticked.

First, we present two simple queries which will bring us to the complete one resulting from their union.

```{r}
#When are reordered products added to the cart
dbGetQuery(sc, 
"SELECT order_id
,   add_to_cart_order
,   SUM(reordered) OVER (PARTITION by order_id ORDER BY add_to_cart_order) AS n_reordered_prod
FROM order_products__prior_tbl
WHERE order_id < 10
")

#Who are regular users
dbGetQuery(sc, 
"SELECT order_id
FROM orders_tbl
WHERE days_since_prior_order < 4
LIMIT 30
")
```

*When do frequent users add reordered products to the cart*

That is, in which order do users who buy on a regular basis (who shop at least twice a week) add reordered products into their carts.

```{r}
dbGetQuery(sc, 
"SELECT b.order_id
,   product_id
,   add_to_cart_order
,   SUM(reordered) OVER (PARTITION by b.order_id ORDER BY add_to_cart_order) AS n_reordered_prod
,   days_since_prior_order
FROM (SELECT order_id, days_since_prior_order
      FROM orders_tbl
      WHERE days_since_prior_order < 4
      LIMIT 30) a
  LEFT JOIN order_products__prior_tbl b
  ON a.order_id=b.order_id
WHERE b.order_id IS NOT NULL
")
```

As we can see, there are some consumers who do not reorder the same products from one day to another. They might be alternating their carts to finish the week, as we suggested, on a balanced diet.

But, on the other hand, most of this users rebuy exactly the same things, and it is precisely these reordered products the ones which are first added to their carts. This corresponds to a very structured shopper who has a clear system and knows exactly what he/she's going to add to the cart before even opening the app. At the end of their shopping lists they might be adding some extra products which they were not predicting or that have been suggested for them, as we saw happens with the latter added items.

Finally, we might be interested in analysing which products are more frequently bought by this consumers' group.

**Regular consumers most ordered products**

```{r}
regular_consumer_txt <- "
SELECT ab.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM (SELECT order_id 
          FROM orders_tbl
          WHERE days_since_prior_order < 4) a
      LEFT JOIN order_products__prior_tbl b
      ON a.order_id = b.order_id
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) ab
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) c
ON ab.product_id = c.product_id
"

(regular_consumer <-
  dbGetQuery(sc, regular_consumer_txt))
```

As we can see, again, their most ordered products do not difer from those most generally ordered. In fact, we could think that these most regular consumers are the ones that end up determining which are the most ordered products among the whole dataset.


#Recommender and Predictions

```{r}
order_products__prior %>%
  select(order_id, product_id) %>%
  left_join(orders, by="order_id") %>%
  filter(user_id <= 10) %>% #d'entrada ho fem només per a 10 usuaris, es recomana anar augmentant la mida poc a poc
  select(product_id, user_id) %>%
  group_by(user_id, product_id) %>%
  summarise(rating = n()) %>%
  rename(user = user_id) %>%
  mutate(item=product_id) %>%
  select(user, item, rating) ->
  user_item_rating
```


```{r}
user_item_rating
```

```{r}
explicit_model <- ml_als_factorization( user_item_rating, iter.max = 5, regularization.parameter = 0.01)
```

```{r}
explicit_model
```

```{r}
order_products__prior %>%
  colnames %>%
  print

orders %>%
  colnames %>%
  print
```

**Matriu A**

```{r}
U <- as.matrix(explicit_model$user.factors[,-1])
V <- as.matrix(explicit_model$item.factors[,-1])
A2 <- U%*%t(V)

library(whisker)
library(magrittr)

i_max <- apply(A2,1,max)
which(A2[6,]==i_max[6])

for(i in 1:nrow(A2)){
  cat(which(A2[i,]==i_max[i]), "\n")
}

recom_prod_txt <- "which(A2[{{userId}},]==i_max[{{userID}}])"

data <- list(userId=1)

recom_prod_txt %>%
  whisker.render(data) %>%
  cat
```

Són indexos dins de la matriu, no productes

per users_factor, producte escalar amb items_factor per columnes, per saber quin és











