---
title: "Final Project"
author: "Laia Esquerrà"
date: "June 30, 2017"
output: html_document
---

Correu: `bartekskorulski@gmail.com`


```{r}
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)
```

- Last hour purchases

```{r}
last_hour_txt <- "
SELECT ab.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM (SELECT order_id 
          FROM orders_tbl
          WHERE order_hour_of_day > 20) a
      LEFT JOIN order_products__prior_tbl b
      ON a.order_id = b.order_id
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) ab
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) c
ON ab.product_id = c.product_id
"

last_hour <-
  dbGetQuery(sc, last_hour_txt)

last_hour
```

- Are the last added products unnecessary?

```{r}
prods_basket_txt <- "
SELECT *
FROM
  (SELECT order_id, MAX (add_to_cart_order) as nmax
  FROM order_products__prior_tbl
  GROUP BY order_id
  LIMIT 10) AS a
    LEFT JOIN order_products__prior_tbl AS b
    ON a.nmax=b.add_to_cart_order
WHERE a.order_id=b.order_id
LIMIT 10
"

(prods_basket <-
  dbGetQuery(sc, prods_basket_txt))
```

- Does a person who buys everyday repeat products?

```{r}
dbGetQuery(sc, 
"
SELECT user_id
,   order_id
,   order_number
,   days_since_prior_order
,   IF(days_since_first_order IS NULL, 0, days_since_first_order) as days_since_first_order
FROM(
SELECT user_id
,   order_id
,   order_number
,   days_since_prior_order
,   SUM(days_since_prior_order) OVER (partition by user_id ORDER BY order_number) AS days_since_first_order
FROM orders_tbl
WHERE user_id <= 10 
) a
ORDER BY user_id
")

dbGetQuery(sc, 
"
SELECT user_id
,   order_id
,   order_number
,   days_since_prior_order
,   IF(days_since_first_order IS NULL, 0, days_since_first_order) as days_since_first_order
FROM(
SELECT order_id
,   SUM(reordered) OVER (PARTITION by order_id ORDER BY product_id) AS n_reordered_prod
FROM order_products__prior_tbl
WHERE order_id = 1 
) a
ORDER BY user_id
")

#When are reordered products added to the cart
dbGetQuery(sc, 
"SELECT order_id
,   add_to_cart_order
,   SUM(reordered) OVER (PARTITION by order_id ORDER BY add_to_cart_order) AS n_reordered_prod
FROM order_products__prior_tbl
WHERE order_id < 10
")

#Who are regular users
dbGetQuery(sc, 
"
SELECT order_id
FROM orders_tbl
WHERE days_since_prior_order < 7
LIMIT 30
")
```

*When do frequent users add reordered products to the cart*

That is, in which order do users who buy on a regular basis (who go at least twice a week) add reordered products into their carts.

```{r}
dbGetQuery(sc, 
"SELECT b.order_id
,   days_since_prior_order
,   add_to_cart_order
,   SUM(reordered) OVER (PARTITION by b.order_id ORDER BY add_to_cart_order) AS n_reordered_prod
FROM (SELECT order_id, days_since_prior_order
      FROM orders_tbl
      WHERE days_since_prior_order < 4
      LIMIT 30) a
  LEFT JOIN order_products__prior_tbl b
  ON a.order_id=b.order_id
")
```

```{r}
dbGetQuery(sc, 
"
SELECT ab.*, c.product_name
FROM(
  SELECT b.order_id
  ,   product_id
  ,   days_since_prior_order
  ,   add_to_cart_order
  ,   SUM(reordered) OVER (PARTITION by b.order_id ORDER BY add_to_cart_order) AS n_reordered_prod
  FROM (SELECT order_id, days_since_prior_order
        FROM orders_tbl
        WHERE days_since_prior_order < 4
        LIMIT 30) a
    LEFT JOIN order_products__prior_tbl b
    ON a.order_id=b.order_id
) ab LEFT JOIN products_tbl c
     ON ab.product_id=c.product_id
")
```



- Regular consumers most ordered products

```{r}
regular_consumer_txt <- "
SELECT ab.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM (SELECT order_id 
          FROM orders_tbl
          WHERE days_since_prior_order < 4) a
      LEFT JOIN order_products__prior_tbl b
      ON a.order_id = b.order_id
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) ab
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) c
ON ab.product_id = c.product_id
"

(regular_consumer <-
  dbGetQuery(sc, regular_consumer_txt))
```














